---
title: 'Data Analysis Assignment #2'
author: "Adams, Dalya"
output:
  html_document: default
---

```{r setup, include=FALSE}
# DO NOT ADD OR REVISE CODE HERE
knitr::opts_chunk$set(echo = FALSE)

# For students:  a = TRUE, b = FALSE
# To knit instructions:  a = FALSE, b = TRUE
a <- TRUE
b <- FALSE
```

### Instructions

R markdown is a plain-text file format for integrating text and R code, and creating transparent, reproducible and interactive reports. An R markdown file (.Rmd) contains metadata, markdown and R code "chunks", and can be "knit" into numerous output types. Answer the test questions by adding R code to the fenced code areas below each item. There are questions that require a written answer that also need to be answered. Enter your comments in the space provided as shown below:

***Answer: (Enter your answer here.)    *** 

Once completed, you will "knit" and submit the resulting .html document and the .Rmd file. The .html will present the output of your R code and your written answers, but your R code will not appear.  Your R code will appear in the .Rmd file. The resulting .html document will be graded and a feedback report returned with comments.  Points assigned to each item appear in the template.

**Before proceeding, look to the top of the .Rmd for the (YAML) metadata block, where the *title*, *author* and *output* are given. Please change *author* to include your name, with the format 'lastName, firstName.'  Do not change the statement a <- TRUE. **

If you encounter issues with knitting the .html, please send an email via Canvas to your TA.

Each code chunk is delineated by six (6) backticks; three (3) at the start and three (3) at the end. After the opening ticks, arguments are passed to the code chunk and in curly brackets. **Please do not add or remove backticks, or modify the arguments or values inside the curly brackets**. An example code chunk is included here: 

```{r exampleCodeChunk, eval = FALSE, echo = TRUE}
# Comments are included in each code chunk, simply as prompts

#...R code placed here

#...R code placed here

```

R code only needs to be added inside the code chunks for each assignment item. However, there are questions that follow many assignment items. Enter your answers in the space provided. An example showing how to use the template and respond to a question follows.

-----

####Submit both the .Rmd and .html files for grading. You may remove the instructions and example problem above, but do not remove the YAML metadata block or the first, "setup" code chunk.  Address the steps that appear below and answer all the questions.  (75 points possible)

-----

##Data Analysis #2

```{r analysis_setup1, eval = a, echo = b, message = FALSE, warning = FALSE}

# Perform the following steps to start the assignment.
 
# 1) Install the following packages from CRAN:  flux, ggplot2, gridExtra, moments, rockchalk, car.

library(flux)
library(ggplot2)
library(gridExtra)
library(rockchalk)
library(car)

# 2) Use the "mydata.csv" file from Assignment #1 or use the file posted on the course site.  Reading
# the files into R will require sep = "" or sep = " " to format data properly.  Use str() to check file
# structure.
library(readr)
mydata <- read_csv("~/Desktop/mydata.csv")
# mydata <- read.csv(file.path("c:..., "mydata.csv"), sep = ",")
str(mydata)

mydata$SEX<-as.factor(mydata$SEX)
mydata$CLASS<-as.factor(mydata$CLASS)

```

-----

(1)(a) (1 point) Form a histogram and QQ plot using RATIO. Calculate skewness and kurtosis using 'rockchalk.' Be aware that with 'rockchalk', the kurtosis value has 3.0 subtracted from it which differs from the 'moments' package. 

```{r Part_1a, eval = a, echo = FALSE}

par(mfrow=c(1,2))

hist(mydata$RATIO, col = "red", main = "Histogram of Ratio", xlab = "Ratio")
qqnorm(mydata$RATIO, col = "red", main = "QQ Plot of Ratio")
qqline(mydata$RATIO, col = "green", lty = 2, lwd = 2)

skewness(mydata$RATIO)
kurtosis(mydata$RATIO)

```

(1)(b) (2 points) Tranform RATIO using log10() to create L_RATIO (see Kabacoff Section 8.5.2, p. 199-200). Form a histogram and QQ plot using L_RATIO. Calculate the skewness and kurtosis. Create a display of six boxplots of L_RATIO differentiated by CLASS.

```{r Part_1b, eval = a, echo = FALSE}

L_RATIO<- log10(mydata$RATIO)

par(mfrow=c(1,2))

hist(L_RATIO, col = "blue", main = "Histogram of Log Ratio", xlab = "L_Ratio")
qqnorm(L_RATIO, col = "blue", main = "QQ Plot of L_RATIO")
qqline(L_RATIO, col = "green", lty = 2, lwd = 2)

skewness(L_RATIO)
kurtosis(L_RATIO)

par(mfrow=c(1,1))
boxplot(L_RATIO~mydata$CLASS, col=c("blue", "purple", "red","pink", "Orange"), main = "Boxplot of L_Ratio across class")


```

(1)(c) (1 point) Test the homogeneity of variance across classes using the bartlett.test() (see Kabacoff Section 9.2.2, p. 222). 

```{r Part_1c, eval = a, echo = FALSE}

bartlett.test(mydata$RATIO ~ mydata$CLASS, data = mydata)
bartlett.test(L_RATIO ~ mydata$CLASS, data = mydata)

```

**Question (2 points):  Based on steps 1.a, 1.b and 1.c, which variable RATIO or L_RATIO exhibits better conformance to a normal distribution with homogeneous variances across age classes?  Why?** 

***Answer: We reject the null hypothesis that the variances are homogenous for the non transformed data. For the log transformed data, we fail to reject the null hypothesis that the variance is homogenous. Visually, viewing the Ratio QQ plot, we see that the data exhibits signs of fat tails, with the divergence from the normal distribution at the tails. This effect is minimized significantly once we transform the data using log10. 
The log transformed data exhibits better conformance to the normal distribution. Despite this, we still see extreme outliers in the boxplots, specifically Class A3. ***

-----

(2)(a) (2 points) Perform an analysis of variance with aov() on L_RATIO using CLASS and SEX as the independent variables (see Kabacoff chapter 9, p. 212-229). Assume equal variances. Peform two analyses. First, fit a model with the interaction term CLASS:SEX. Then, fit a model without CLASS:SEX. Use summary() to obtain the analysis of variance tables (Kabacoff chapter 9, p. 227).

```{r Part_2a, eval = a, echo = FALSE}

nointer<-aov(L_RATIO~mydata$CLASS+mydata$SEX)
inter<-aov(L_RATIO~mydata$CLASS+mydata$SEX+mydata$CLASS*mydata$SEX)

summary(inter)
summary(nointer)

```

**Question (2 points):  Compare the two analyses.  What does the non-significant interaction term suggest about the relationship between L_RATIO and the factors CLASS and SEX?**

***Answer: The lack of significance of the interaction terms suggests that these two variables do not cross or interact with each other. Class does not influence sex and sex does not influence class. They are independent of each other. If we were to graph sex and class, these lines would be parallel and never cross or interact with each other. Since Class and Sex are both significant in this model, we can drop the interaction term since there is no relationship/interaction being missed by dropping the interaction term.   ***

(2)(b) (2 points) For the model without CLASS:SEX (i.e. an interaction term), obtain multiple comparisons with the TukeyHSD() function. Interpret the results at the 95% confidence level (TukeyHSD() will adjust for unequal sample sizes). 

```{r Part_2b, eval = a, echo = FALSE}
TukeyHSD(nointer, conf.level = 0.95)

```

**Question (2 points) :  Interpret the trend across classes. Do these results suggest male and female abalones can be combined into a single category labeled as 'adults?' If not, why not?**

***Answer: At a 5% level of significance, we reject the null hypothesis that the difference of means for all variable combinations is 0, except A2-A1 and M-F. These results suggest that M-F and A2-A1 can be combined. In the case of M-F, they can be combined to create a new variable, Adult. A2-A1, the combination is not as simple, since infant already exists and these classes likely correlate with the sex classification Infant.    ***

-----

(3)(a) (2 points) Use combineLevels() from the 'rockchalk' package to combine "M" and "F" into a new level, "ADULT". This will necessitate defining a new variable, TYPE, in mydata which will have two levels:  "I" and "ADULT". Use par() to form two histograms of VOLUME. One should display infant volumes, and the other:  adult volumes. 

```{r Part_3a, eval = a, echo = b}

mydata$TYPE <- combineLevels(mydata$SEX, levs = c("F","M"), "ADULT")

summary(mydata)
par(mfrow=c(1,2))
inf.volume<-subset(mydata$VOLUME, mydata$TYPE=="I")
adult.volume<-subset(mydata$VOLUME, mydata$TYPE=="ADULT")

hist(inf.volume, main= "Infant Volume", col = "pink")
hist(adult.volume, main = "Adult Volume", col = "red")

```

**Question (2 points): Compare the histograms.  How do the distributions differ? What are the implications in terms of identifying and, ultimately, separating infants from adults based on VOLUME?**

***Answer: Infant Abalones volumes are grouped closer to the left tail of the histogram. This indicates that the weights of abalones categorized as infants builds up between 0 to 200. The volume of adult abalone is higher than that of infants. The majority of adult volumes are between 300 and 600. While it is clear there is a distinction in the average volume of an adult abalone and an infant abalone, there is still a high frequency of adult abalone with weights common for infants and infant abalone with weights common for adult abalone. This will likely mean there isn't a clear-cut way to separate infant and adult abalone based on volume. Using volume as the deciding factor for harvesting abalone or not will always result in some infants being harvested or some adults not being harvested.   ***

(3)(b) (3 points) Create a scatterplot of SHUCK versus VOLUME and a scatterplot of their base ten logarithms, labeling the variables as L_SHUCK and L_VOLUME. Please be aware the variables, L_SHUCK and L_VOLUME, present the data as orders of magnitude (i.e. VOLUME = 100 = 10^2 becomes L_VOLUME = 2). Use color to differentiate CLASS in the plots. Repeat using color to differentiate only by TYPE. 

```{r Part_3b, eval = a, echo = FALSE}

L_SHUCK<-log10(mydata$SHUCK)
L_VOLUME<-log10(mydata$VOLUME)

#By Class
par(mfrow=c(1,2))
with(mydata, plot(mydata$SHUCK, mydata$VOLUME, col=mydata$CLASS, pch=20, xlab="Shuck", ylab = "Volume", main = "Shuck and Volume by Class"))
legend("bottomright",inset = .02, c("A1","A2","A3","A4","A5"), fill = c("black", "red", "green", "blue", "skyblue"),cex = 0.8)

with(mydata, plot(L_SHUCK, L_VOLUME, col=mydata$CLASS, pch=20, main = "L_Shuck and L_Volume by Class"))
legend("bottomright", inset = .02, c("A1","A2","A3","A4","A5"), fill = c("black", "red", "green", "blue", "skyblue"), cex = 0.8)

#by TYPE
par(mfrow=c(1,2))
with(mydata, plot(mydata$SHUCK, mydata$VOLUME, col=mydata$TYPE, pch=20, xlab="Shuck", ylab = "Volume", main = "Shuck and Volume by Type"))
legend("bottomright",inset = .02, c("INFANT","ADULT"), fill = c("black", "red"),cex = 0.8)

with(mydata, plot(L_SHUCK, L_VOLUME, col=mydata$TYPE, pch=20, main = "L_Shuck and L_Volume by Type"))
legend("bottomright",inset = .02, c("INFANT","ADULT"), fill = c("black", "red"),cex = 0.8)
```

**Question (3 points):  Compare the two scatterplots. What effect(s) does log-transformation appear to have on any relationship between SHUCK weight and VOLUME? Where do the various CLASS levels appear in the plots? Where do the levels of TYPE appear in the plots?**

***Answer: Prior to the log transformation, shuck and volume exhibited a strong positive linear correlation. After the log transformation this positive linear relationship was still present. Something that the log transformation appears to have done is separate the infant data points from the adult data points. In both scatterplots for the non-transformed data, all classes and types are grouped together and there is not a clear distinction between the different classes or types. Once the log transformation is applied, Class A1 shows as a distinct class, in black and the separation between infant and adult also becomes clear, with infant being black and adult being red. ***

-----

(4)(a) (3 points) Since abalone growth slows after class A3, infants in classes A4 and A5 are considered mature and candidates for harvest. Reclassify the infants in classes A4 and A5 as ADULTS. This reclassification can be achieved using combineLevels(), but only on the abalones in classes A4 and A5.  You will use this recoded TYPE variable, in which the infants in A4 and A5 were
reclassified as ADULTS, for the remainder of this data analysis assignment.

Regress L_SHUCK as the dependent variable on L_VOLUME, CLASS and TYPE (see Kabacoff Section 8.2.4, p. 178-186, the Data Analysis Video #2 and Black Section 14.2). Use the multiple regression model: L_SHUCK ~ L_VOLUME + CLASS + TYPE. Apply summary() to the model object to produce results.

```{r Part_4a, eval = a, echo = FALSE}

mydata$TYPE[mydata$CLASS=="A4"|mydata$CLASS=="A5"]<-combineLevels(mydata$TYPE[mydata$CLASS=="A4"|mydata$CLASS=="A5"], levs = c("I","ADULT"),"ADULT")

regression<-lm(L_SHUCK~L_VOLUME+CLASS+TYPE, data = mydata)
summary(regression)
```

**Question (2 points):  Interpret the trend in coefficient estimates for CLASS levels (Hint:  this question is not asking if the estimates are statistically significant. It is asking for an interpretation of the pattern in these coefficients, and how this pattern relates to the earlier displays).**

***Answer: When looking at the variables for Class in the multiple linear regression model,  we notice that as the classes increase from A3 to A5, the negative impact increases, or the coefficient becomes increasingly negative, from -0.047 to -0.117. In our coefficients, we see being Class A3, results in a 4.7% decrease in log transformed Shuck weight, being Class A4 results in a 7.6% decrease in log transformed Shuck weight and at Class A5, the log transformed Shuck weight decreases by 12%. We see that Adult results in a 2% increase of log transformed. ***

**Question (2 points):  Is TYPE an important predictor in this regression? (Hint:  This question is not asking if TYPE is statistically significant, but rather how it compares to the other independent variables in terms of its contribution to predictions of L_SHUCK.)  Explain your conclusion.**

***Answer: Type is significant in that being classified as ADULT results in a 2% increase in the log shuck. This fits with what we’d expect, an adult abalone would weigh more than an infant abalone. Our Class variables did not reinforce this idea, since the adult classifications of A4 and A5 resulted in a decrease in the log shuck weight. While this 2% increase is statistically significant and fits with what we would expect, a 2% increase is not very materially significant, especially considering the strong material significance of L_Volume in relation to L_Shuck.   ***

-----

The next two analysis steps involve an analysis of the residuals resulting from the regression model in (4)(b) (see Kabacoff Section 8.2.4, p. 178-186, the Data Analysis Video #2).

-----

(5)(a) (3 points) If "model" is the regression object, use model$residuals and construct a histogram and QQ plot. Compute the skewness and kurtosis. Be aware that with 'rockchalk,' the kurtosis value has 3.0 subtracted from it which differs from the 'moments' package. 

```{r Part_5a, eval = a, echo = FALSE}
reg.resid<-regression$residuals

par(mfrow=c(1,2))
hist(reg.resid, main = "Residuals", col ="darkblue", xlab = "Regression Residuals" )

qqnorm(reg.resid, col = "darkblue", main = "QQ Plot of Residuals")
qqline(reg.resid, col = "green", lty = 2, lwd = 2)

skewness(reg.resid)
kurtosis(reg.resid)
```

(5)(b) (3 points) Plot the residuals versus L_VOLUME coloring the data points by CLASS, and a second time coloring the data points by TYPE (Keep in mind the y-axis and x-axis may be disproportionate which will amplify the variability in the residuals). Present boxplots of the residuals differentiated by CLASS and TYPE (These four plots can be conveniently presented on one page using par(mfrow..) or grid.arrange(). Test the homogeneity of variance of the residuals across classes using the bartlett.test() (see Kabacoff Section 9.3.2, p. 222).  

```{r Part_5b, eval = a, echo = FALSE}
par(mfrow=c(2,2))

#Colored by Class
with(mydata, plot(L_VOLUME, reg.resid, col=mydata$CLASS, pch=20, xlim = c(0,3.5)))
legend("bottomleft", inset = .02, c("A1","A2","A3","A4","A5"), fill = c("black", "red", "green", "blue", "skyblue"), cex = 0.8)

#Colored by Type
with(mydata, plot(L_VOLUME, reg.resid, col=mydata$TYPE, pch=20, xlim = c(0,3.5)))
legend("bottomleft",inset = .02, c("INFANT","ADULT"), fill = c("black", "red"),cex = 0.8)

#Boxplots by Class
boxplot(reg.resid~mydata$CLASS, col=c("blue", "purple", "red","pink", "Orange"), main = "Residuals across class")

#Boxplots by Type
boxplot(reg.resid~mydata$TYPE, col=c("blue", "red"), main = "Residuals across type")

bartlett.test(reg.resid~mydata$CLASS, data = mydata)
```

**Question (3 points):  What is revealed by the displays and calculations in (5)(a) and (5)(b)?  Does the model 'fit'?  Does this analysis indicate that L_VOLUME might be useful for harvesting decisions? Discuss. **  

***Answer: When viewing the skewness and kurtosis of the residuals, we see they are close to normal, but has a slightly higher than normal kurtosis, meaning there are likely outliers in the tail. This is shown in the QQ-Plot and very clearly in the residuals by Type box plot. The Bartlett Test fails to reject the null hypothesis that the variance is homogenous. This would indicate that our residuals have a variance of 0. Despite this, the scatterplot of the residuals versus log volume appear to be coned shaped, with more points to the right. This does not indicate that the residuals are homoskedastic, but instead that this model suffers from heteroskedasticity. This leads us to believe that L_Volume is not useful in harvesting decisions, as the assumption that our residual variance is equal to 0 would be violated.***

-----

There is a tradeoff faced in managing abalone harvest. The infant population must be protected since it represents future harvests. On the other hand, the harvest should be designed to be efficient with a yield to justify the effort. This assignment will use VOLUME to form binary decision rules to guide harvesting. If VOLUME is below a "cutoff" (i.e. specified volume), that individual will not be harvested. If above, it will be harvested. Different rules are possible.

The next steps in the assignment will require plotting of infants versus adults. For this  plotting to be accomplished, similar "for loops" must be used to compute the harvest proportions. These loops must use the same value for the constants min.v and delta; and, use the same statement "for(k in 1:1000)."  Otherwise, the resulting infant and adult proportions cannot be directly  compared and plotted as requested. Note the example code supplied below.

-----

(6)(a) (2 points) Calculate the proportion of infant and adult abalones which fall beneath a specified volume or "cutoff." A series of volumes covering the range from minimum to maximum abalone volume will be used in a "for loop" to determine how the harvest proportions change as the "cutoff" changes. Example code for doing this is provided.

```{r Part_6a, eval = a, echo = b}

idxi <- mydata$TYPE == "I"
idxa <- mydata$TYPE == "ADULT"

max.v <- max(mydata$VOLUME)
min.v <- min(mydata$VOLUME)
delta <- (max.v - min.v)/1000
prop.infants <- numeric(0)
prop.adults <- numeric(0)
volume.value <- numeric(0)

total.infants <- sum(idxi)  
total.adults <- sum(idxa)

for (k in 1:1000) { 
	value <- min.v + k*delta
	volume.value[k] <- value
	prop.infants[k] <- sum(mydata$VOLUME[idxi] <= value)/total.infants
	prop.adults[k] <-  sum(mydata$VOLUME[idxa] <= value)/total.adults
}

# prop.infants shows the impact of increasing the volume cutoff for
# harvesting. The following code shows how to "split" the population at
# a 50% harvest of infants.

n.infants <- sum(prop.infants <= 0.5)
split.infants <- min.v + (n.infants + 0.5)*delta  # This estimates the desired volume.

n.adults <- sum(prop.adults <= 0.5)
split.adults <- min.v + (n.adults + 0.5)*delta

```

(6)(b) (2 points) Present a plot showing the infant proportions and the adult proportions versus volume. Compute the 50% "split" volume.value for each and show on the plot.   

```{r Part_6b, eval = a, echo = FALSE}
split.infants
split.adults

plot(volume.value, prop.infants, col="red",type="l", xlab = "Volume", ylab = "Proportion", main = "Proportion of Adults and Infant Protected")
lines(volume.value, prop.adults, col="blue")
abline(v=split.infants)
abline(v=split.adults)
abline(h=0.5)
text(170,0.5, "134.07", pos = 1, cex = .6)
text(420, 0.5, "384.07", pos = 1, cex = .6)
legend("bottomright", c("INFANT", "ADULT"), fill = c("red", "blue"), cex = 0.8)

```

**Question (2 points):  The two 50% "split" values serve a descriptive purpose illustrating the difference between the populations. What do these values suggest regarding possible cutoffs for harvesting?** 

***Answer: Viewing these two 50% split values, we can see what is potentially lost and gained by setting a volume cutoff. If we were to select the 50% adult split value, we would harvest an extremely small number of infant abalone but we are also losing out on 50% of the adult abalone population. If we were to set the cut off at he 50% infant split value, we are able to harvest almost all of the adult abalone, but we are also harvesting 50% of the infant abalone population, which is not the goal if we want sustainable abalone harvesting.  All in all, these values suggest that there are trade offs depending on which value is selected. In order to choose the correct one, the risk vs. reward must be established. If harvesting an infant abalone is very bad for the population, than missing 50% of the adult abalone may be worth it. If abalone reproduce in large numbers and grow to maturity quickly, maybe harvesting 50% of the infant abalone population is not harmful. These are the kind of things that need to be taken into consideration before choosing a volume threshold.    ***

-----

This part will address the determination of a volume.value corresponding to the observed maximum difference in harvest percentages of adults and infants. To calculate this result, the proportions from (6) must be used. These proportions must be converted from "not harvested" to "harvested" proportions by using (1 - prop.infants) for infants, and (1 - prop.adults) for adults. The reason the proportion for infants drops sooner than adults is that infants are maturing and becoming adults with larger volumes.

-----

(7)(a) (1 point) Evaluate a plot of the difference ((1 - prop.adults) - (1 - prop.infants)) versus volume.value. Compare to the 50% "split" points determined in (6)(a). There is considerable variability present in the peak area of this plot. The observed "peak" difference may not be the best representation of the data. One solution is to smooth the data to determine a more representative estimate of the maximum difference.

```{r Part_7a, eval = a, echo = FALSE}

difference<-((1 - prop.adults) - (1 - prop.infants))
plot(volume.value, difference, col="blue", type = "l", xlab = "Volume", ylab = "Difference in Proportions Harvested", main = "Difference in Harvest Proportions")

```

(7)(b) (1 point) Since curve smoothing is not studied in this course, code is supplied below. Execute the following code to determine a smoothed version of the plot in (a). The procedure is to individually smooth (1-prop.adults) and (1-prop.infants) before determining an estimate of the maximum difference. 

```{r Part_7b, eval = a, echo = b}

y.loess.a <- loess(1 - prop.adults ~ volume.value, span = 0.25,
	family = c("symmetric"))
y.loess.i <- loess(1 - prop.infants ~ volume.value, span = 0.25,
	family = c("symmetric"))
smooth.difference <- predict(y.loess.a) - predict(y.loess.i)

```

(7)(c) (3 points) Present a plot of the difference ((1 - prop.adults) - (1 - prop.infants)) versus volume.value with the variable smooth.difference superimposed. Determine the volume.value corresponding to the maximum of the variable  smooth.difference (Hint:  use which.max()).Show the estimated peak location corresponding to the cutoff determined.

```{r Part_7c, eval = a, echo = FALSE}

plot(volume.value, difference, col="blue", type = "l", xlab = "Volume", ylab = "Difference in Proportions Harvested", main = "Difference in Harvest Proportions")
lines(volume.value, smooth.difference, lty=2)
abline(v=261.548, lty=3)
text(270, 0.0, "Volume = 261.548", pos = 4, srt=90)

Max.Diff.value<-volume.value[which.max(smooth.difference)]
Max.Diff.value
```

(7)(d) (1 point) What separate harvest proportions for infants and adults would result if this cutoff is used? (NOTE:  the adult harvest proportion is the "true positive rate" and the infant harvest proportion is the "false positive rate.")

Code for calculating the adult harvest proportion is provided.

*(1-prop.adults)[which.max(smooth.difference)]  # [1] 0.7416332*

```{r Part_7d, eval = a, echo = FALSE}

max.diff.adult<-(1-prop.adults)[which.max(smooth.difference)] 
max.diff.adult
max.diff.infant<-(1-prop.infants)[which.max(smooth.difference)] 
max.diff.infant
```

-----

There are alternative ways to determine cutoffs. Two such cutoffs are described below.

-----

(8)(a) (2 points) Harvesting of infants in CLASS "A1" must be minimized. The smallest volume.value cutoff that produces a zero harvest of infants from CLASS "A1" may be used as a baseline for comparison with larger cutoffs. Any smaller cutoff would result in harvesting infants from CLASS "A1."  

Compute this cutoff, and the proportions of infants and adults with VOLUME exceeding this cutoff. Code for determining this cutoff is provided.

*volume.value[volume.value > max(mydata[mydata\$CLASS == "A1" &*
  *mydata\$TYPE == "I", "VOLUME"])][1] # [1] 206.9844*

```{r Part_8a, eval = a, echo = FALSE}

zero.a1<-volume.value[volume.value > max(mydata[mydata$CLASS == "A1" & mydata$TYPE == "I", "VOLUME"])][1]
zero.a1

zero.a1.adult<-sum(mydata[mydata$TYPE=="ADULT", "VOLUME"]>volume.value[volume.value > max(mydata[mydata$CLASS == "A1" & mydata$TYPE == "I", "VOLUME"])][1])/sum(mydata$TYPE=="ADULT")
zero.a1.adult

zero.a1.infant<-sum(mydata[mydata$TYPE=="I", "VOLUME"]>volume.value[volume.value > max(mydata[mydata$CLASS == "A1" & mydata$TYPE == "I", "VOLUME"])][1])/sum(mydata$TYPE=="I")
zero.a1.infant
```

(8)(b) (2 points) Another cutoff can be determined for which the proportion of adults not harvested equals the proportion of infants harvested. This cutoff would equate these rates; effectively, our two errors:  'missed' adults and wrongly-harvested infants. This leaves for discussion which is a greater loss:  a larger proportion of adults not harvested or infants harvested?  This cutoff is 237.7383. Calculate the separate harvest proportions for infants and adults using this cutoff. Code for determining this cutoff is provided.

*volume.value[which.min(abs(prop.adults - (1-prop.infants)))] # [1] 237.7383*

```{r Part_8b, eval = a, echo = FALSE}

equal<-volume.value[which.min(abs(prop.adults - (1-prop.infants)))]
equal

equal.adult<-sum(mydata[mydata$TYPE=="ADULT", "VOLUME"]>volume.value[which.min(abs(prop.adults - (1-prop.infants)))])/sum(mydata$TYPE=="ADULT")
equal.adult

equal.infant<-sum(mydata[mydata$TYPE=="I", "VOLUME"]>volume.value[which.min(abs(prop.adults - (1-prop.infants)))])/sum(mydata$TYPE=="I")
equal.infant
```

-----

(9)(a) (7 points) Construct an ROC curve by plotting (1 - prop.adults) versus (1 - prop.infants). Each point which appears corresponds to a particular volume.value. Show the location of the cutoffs determined in (7) and (8) on this plot and label each. 

```{r Part_9, eval = a, echo = FALSE}
d.prop.adults<-(1-prop.adults)
d.prop.infants<-(1-prop.infants)

plot(d.prop.infants, d.prop.adults, col="blue", type = "l", xlab = "Infant Harvest Proportions", ylab = "Adult Harvest Proportions", main = "ROC Curve of adult and infant harvest proportions")
abline(0,1, lty=2)

points(.17, .74, pch=1)
text(.25,.74,"Max. Difference Vol =261.5", pos=1, cex = .5)

points(.29, .83, pch=1)
text(.29,.85, "Zero A1 inf. Vol =207", pos=3, cex = .5)

points(.22, .78, pch=1)
text(.22, .75, "Equal harvest/conserve vol =237.7", pos=4, cex = .5)

```

(9)(b) (1 point) Numerically integrate the area under the ROC curve and report your result. This is most easily done with the auc() function from the "flux" package.   Areas-under-curve, or AUCs, greater than 0.8 are taken to indicate good discrimination potential. 

```{r Part_9b, eval = a, echo = FALSE}

auc(d.prop.infants, d.prop.adults)

```

-----

(10)(a) (3 points) Prepare a table showing each cutoff along with the following:
 	1) true positive rate (1-prop.adults,
 	2) false positive rate (1-prop.infants),
 	3) harvest proportion of the total population
 	
```{r Part_10, eval = a, echo = FALSE} 	
tot1<-subset(mydata$VOLUME, mydata$VOLUME>261.5478)
ltot1<-length(tot1)
tot.len<-length(mydata$VOLUME)
PY1<-ltot1/tot.len

tot2<-subset(mydata$VOLUME, mydata$VOLUME>206.9844)
ltot2<-length(tot2)
PY2<-ltot2/tot.len

tot3<-subset(mydata$VOLUME, mydata$VOLUME>237.7383)
ltot3<-length(tot3)
PY3<-ltot3/tot.len


max.difference<-c(Max.Diff.value, max.diff.adult, max.diff.infant, PY1)
zero.A1.infants<-c(zero.a1,zero.a1.adult, zero.a1.infant, PY2)
equal.error<-c(equal, equal.adult, equal.infant, PY3)

trix<-rbind(max.difference, zero.A1.infants, equal.error)
colnames(trix)<-c("Volume", "TPR", "FPR", "Prop Yield")
trix

```
 	
**Question: (3 points) Based on the ROC curve, it is evident a wide range of possible "cutoffs" exist. Compare and discuss the three cutoffs determined in this assignment. **   

***Answer: The first cutoff at a volume of 262. This cut off is determined by identifying the maximum distance between the infant and adult abalone lines. At this point, the largest numbers of infant abalone are protected and adult abalone are harvested. If harvesting infant abalone is very risky for the population, than this cut off is not appropriate. The second cutoff value is 207. This is the volume criterion set if Infant abalones of Class A1 are never to be harvested. If protecting the infant abalone is of utmost importance, this is an appropriate cutoff to use. The problem is that the number of adult abalone that are not harvested is also high. Using this criterion hinders abalone harvesting, as almost 30% of harvestable abalone are mistakenly categorized as infants. The final criterion is 238. This cutoff forces the adult and infant error equal, or the missed adult abalones equal to the wrongly harvested infant abalone. Again, we are faced with the same question, is missed opportunity more risky or harvesting infant abalone? Depending on how this is answered determines which cutoff should be used. In all the cutoffs discussed, infant abalone will be wrongly harvested and adult abalone will be missed.     ***

-----

**Question (5 points):  Assume you are expected to make a presentation of your analysis to the investigators How would you do so?  Consider the following in your answer:  1) Would you make a specific recommendation or outline various choices and tradeoffs? 2)  What qualifications or limitations would you present regarding your analysis?  3) If it is necessary to proceed based on the current analysis, what suggestions would you have for implementation of a cutoff?  4)  What suggestions would you have for planning future abalone studies of this type? **  

***Answer:In reviewing the results from this analysis, a correct or simple solution that satisfies both requirements, not harvesting infants and harvesting adults, does not present itself. This makes presenting a single and absolute solution a poor choice. Since some infant abalone will be harvested and some adult abalone will not be harvested, the risk to reward needs to be identified before making a decision on which Volume to use as a cutoff when harvesting abalone. Presenting all the options, and the benefits and pitfalls of each option seems like the better choice when presenting the finding of this study. When presenting the options of volume cutoff, I would explain the risk inherent with the true positive versus false positive percentages. I would also explain that the variation present in the data makes providing a clear cutoff difficult. When proceeding with implementing a volume based cutoff, knowing the risk vs reward for each choice is very important, as well as an understanding of the overall goal. If the overall goal is to not harvest infant abalones, than a cutoff can be chosen to support this goal. If the goal is to limit infant harvesting, while maximizing the adult harvest than a different volume cutoff can be selected. Understanding what the overall goal, and the implications of that goal are the most important part in choosing the cutoff based on the analysis provided.  For future abalone studies, the variance present in the class specific data will need to be addressed or accounted for. Is it very common for infant abalone to weigh the same as adult abalone? If so, a better system for distinguishing infant and adult abalone may need to be identified.    ***